A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-10-18 kobakian-cook.Rmd
simplified images

```{r setup}
library(knitcitations)
library(kableExtra)
library(sp)
library(sf)
library(sugarbag)
library(tidyverse)
library(cartogram)
library(ggthemes)
library(cowplot)

@@

# function to allocate colours to regions
aus_colours <- function(sir_p50){
  value <- case_when(
    sir_p50 <  0.74 ~ "#33809d",
    sir_p50 >= 0.74 & sir_p50 < 0.98 ~ "#aec6c7",
    sir_p50 >= 0.98 & sir_p50 < 1.05 ~ "#fff4bc",
    sir_p50 >= 1.05 & sir_p50 < 1.45 ~ "#ff9a64",
    sir_p50 >= 1.45 ~ "#ff3500",
    TRUE ~ "#FFFFFF")
  return(value)
}

@@

```{r thyroiddata}
sa2 <- absmapsdata::sa22011 %>% 
  filter(!st_is_empty(geometry)) %>% 
  filter(!state_name_2011 == "Other Territories") %>% 
  filter(!sa2_name_2011 == "Lord Howe Island")

# sa2 <- sa2 %>% rmapshaper::ms_simplify(keep = 0.05, keep_shapes = TRUE)

SIR <- read_csv("data/SIR Downloadable Data.csv") %>% 
  filter(SA2_name %in% sa2$sa2_name_2011) %>% 
  dplyr::select(Cancer_name, SA2_name, Sex_name, p50) %>% 
  filter(Cancer_name == "Thyroid", Sex_name == "Females")
ERP <- read_csv("data/ERP.csv") %>%
  filter(REGIONTYPE == "SA2", Time == 2011, Region %in% SIR$SA2_name) %>% 
  dplyr::select(Region, Value)

# Alternative maps
# Join with sa2 sf object
sa2thyroid_ERP <- SIR %>% 
  left_join(sa2, ., by = c("sa2_name_2011" = "SA2_name")) %>%
  left_join(., ERP %>% 
              dplyr::select(Region, 
              Population = Value), by = c("sa2_name_2011"= "Region")) %>% 
  filter(!st_is_empty(geometry))

sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  #filter(!is.na(Population)) %>% 
  filter(!sa2_name_2011 == "Lord Howe Island") %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf() 
```

@@

```{r fullhexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}

if (!file.exists("data/aus_hexmap.rda")) {

## Create centroids set
centroids <- sa2 %>% 
  create_centroids(., "sa2_name_2011")

## Create hexagon grid
grid <- create_grid(centroids = centroids,
                    hex_size = 0.2,
                    buffer_dist = 5)

## Allocate polygon centroids to hexagon grid points
aus_hexmap <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.2,
  ## same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities,
  width = 35,
  verbose = FALSE
)

save(aus_hexmap, file = "data/aus_hexmap.rda") }

@@

## Prepare to plot
fort_hex <- fortify_hexagon(data = aus_hexmap,
                            sf_id = "sa2_name_2011",
                            hex_size = 0.2) %>% 
            left_join(sa2thyroid_ERP %>% select(sa2_name_2011, SIR, p50))

fort_aus <- sa2thyroid_ERP %>%
  fortify_sfc()

## Make a plot
aus_hexmap <- ggplot() +
  geom_polygon(aes(x = long,  y = lat,  group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey",  size = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal()

@@

# Contiguous Cartograms

if (!file.exists("data/ncont.rda")) {
cont <- sa2thyroid_ERP %>% 
  mutate(Population = Population + 1) %>% 
  cartogram_cont(., weight = "Population", itermax = 20) %>%
  st_as_sf()

save(cont, file = "data/cont.rda") }

@@

tas_ggcont <- ggplot(cont) + 
  geom_sf(fill = "lightgrey", colour = "black", data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + guides(fill=FALSE) +
  theme_map() + ggtitle("a")
#tas_ggcont

```{r contcart}
# Non - Contiguous Cartograms

if (!file.exists("data/ncont.rda")) {
ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1.5,
                         weight = "Population") %>% 
                          st_as_sf()

save(ncont, file = "data/ncont.rda")
}

```{r plotcontcart}
load("data/ncont.rda")

tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "lightgrey", colour = "black",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  guides(fill=FALSE) +
  theme_map() + ggtitle("b")
#tas_ggncont

```{r dorl}
# Non - Contiguous Dorling Cartograms

dorl <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
tas_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "lightgrey", colour = "black",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +
  guides(fill = FALSE) +
  theme_map() + ggtitle("c")
#tas_ggdorl
```

@@

```{r hexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid for males, they range from much lower than average (blue) to much higher than average (red)."}
aus_hexmap
```

@@

```{r centroids_plot, fig.cap = "The Australian Cancer Atlas data has determined the colour of each Statistical Area of Australian at Level 2. A choropleth map (a) of Standardised Incidence Rates (SIRs) is paired with a hexagon tile map (b) to contrast the colours that are made obvious when every SA2 is equally represented."}
centroids <- create_centroids(
  shp_sf = sa2 %>% filter(state_name_2011 == "Tasmania"), 
  sf_id = "sa2_name_2011")

hex_size <- .2
buffer_dist <- 2

bbox <- tibble(min = c(min(centroids$longitude), min(centroids$latitude)),
  max = c(max(centroids$longitude), max(centroids$latitude)))

grid <- tibble::as_tibble(expand.grid(hex_long = seq(bbox$min[1] - buffer_dist,
  bbox$max[1] + buffer_dist,
  hex_size),
  hex_lat = seq(bbox$min[2] - buffer_dist,
    bbox$max[2] + buffer_dist,
    hex_size)))

@@

tas_centroids <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) + 
  geom_point(aes(x=longitude, y = latitude, colour = SIR), size = 3, shape = 19,
             data = st_drop_geometry(sa2thyroid_ERP) %>% 
               select(sa2_name_2011, SIR) %>%  
               left_join(centroids)) +
  scale_colour_identity() +
  theme_void() + 
  coord_equal()

tas_hexmap <- ggplot() +
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) + 
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR),
               data = fort_hex %>% filter(sa2_name_2011 %in% fort_tas$sa2_name_2011)) +
  scale_fill_identity() +
  theme_void() + 
  coord_equal()

grid.arrange(tas_centroids, tas_hexmap, nrow=1)
```

@@

```{r grid1}
#grid <- create_grid(centroids = centroids, hex_size = 0.2, buffer_dist = 1.2)

g1 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) + 
  geom_point(aes(x=longitude, y = latitude), data = centroids, colour = "#374926") + theme_void() + 
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#64b1be", data = grid, size = 1.5) +
  coord_equal() + ggtitle("a")
```

@@

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is the size specified by `hex_size`. Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. Similarly, the rows are created from the latitude values and the buffer distance. A unique hexagon location is created from all intersections of the longitude columns and latitude rows. Figure \ref{fig:grid2} shows the original grid on the left, to allow for tessellating hexagons, every second latitude row on the grid is shifted right, by half of the hexagon size. The grid for tessellation is shown on the right.

```{r grid2, fig.cap = "Complete set of grid points to create a tile map.", height = 4}
# Find every second latitude
shift_lat <- grid %>% dplyr::select(hex_lat) %>%
  dplyr::distinct() %>%
  dplyr::filter(dplyr::row_number() %% 2 == 1) %>% unlist()

# Shift the longitude of every second latitude to the right to make hex structure
grid <- grid %>%
  dplyr::mutate(hex_long = ifelse(hex_lat %in% shift_lat, hex_long,
    hex_long + (hex_size / 2))) %>%
  dplyr::mutate(id=1:NROW(.))  %>%
  dplyr::mutate(assigned=FALSE)

g2 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) + 
  geom_point(aes(x=longitude, y = latitude), data = centroids, colour = "#374926") + theme_void() + 
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#64b1be", data = grid, size = 1.5) +
  coord_equal() + ggtitle("b")

grid.arrange(g1, g2, nrow = 1)
```

@@

The closest latitude row and longitude column are found for each centroid location. Then rows and columns of centroids are divided into 20 groups. The amount of rows in each latitude group and the amount of columns in each longitude group are used as the width of rolling windows. The rolling windows can be seen on the bottom and right of the grid shown in Figure \ref{fig:grid3}. This will tailor the available grid points to those most likely to be used. It also helps reduce the amount of time taken, as it decreases the amount of points considered for each centroid allocation.

@@

```{r grid3, fig.cap = "The remaining hexagon locations from all the possible points in the initial grid after the buffer is applied. The blue dots show the grid points left to choose from after the buffer step."}

full_grid <- grid <- grid %>%
  mutate(hex_long_int = dense_rank(hex_long)-1,
    hex_lat_int = dense_rank(hex_lat)-1)

nlong <- length(unique(grid$hex_long))
nlat <- length(unique(grid$hex_lat))

  # Add grid row and column values to centroids data set 
  centroids <- centroids %>%
    mutate(
      # long int for integer value of longitude column
      long_int = round((longitude - min(grid$hex_long)) / 
                         (max(grid$hex_long) - min(grid$hex_long)) * nlong, 0),
      # lat int for integer value of latitude column
      lat_int = round((latitude - min(grid$hex_lat)) / 
                        (max(grid$hex_lat) - min(grid$hex_lat)) * nlat, 0)
    )

  # Amount of lats and longs in each group of rows and columns
  lat_size <- round(nlat / 20, 0)
  long_size <- round(nlong / 20, 0)

  # Make a list of the min and max of the groups
  # Effectively creates manual sliding windows
  nlat_list <- purrr::map2(seq(1:nlat), lat_size + seq(1:nlat), c)
  nlong_list <- purrr::map2(seq(1:nlong), long_size + seq(1:nlong), c)

  # LATITUDE ROWS FILTER

@@

    max_int <- min(bounds[2], maximum)

    cents_in <- filter(cents, between(lat_int, bounds[1], max_int))
    return(cents_in)
  }

@@

    while (max_int > maximum) {
      max_int <- max_int - 1
    }

    cents_in <- filter(cents, between(long_int, bounds[1], max_int))
    return(cents_in)
  }

@@

  # DEFINE FUNCTION
  # find the min and max longitude for each latitude
  range_rows <- purrr::map_dfr(
    .x = lat_windows,
    .f = function(data) {
      data %>%
        dplyr::summarise(
          long_min = ifelse(purrr::is_empty(long_int), NA, min(data$long_int)),
          long_max = ifelse(purrr::is_empty(long_int), NA, max(data$long_int))
        )
    }
  )

@@

  # smooth the minimums
  av_range_cols <- purrr::map_dfr(.x = nlong_list, mean_range, range_cols) %>%
    bind_cols(long_id = c(seq(1:nlong) + round(long_size / 2)), .)

  # APPLY A BUFFER
  # change buffer to amount of hexagons (ints) either side
  hex_buffer <- floor(buffer_dist / hex_size)

@@

ggplot() + 
   geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) + 
  geom_point(aes(x=longitude, y = latitude),data = centroids, colour = "#374926") + 
  theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#9ad5c0", data = full_grid, size = 0.5) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#64b1be", data = grid, size = 2)

@@

```{r buffers, fig.cap = "Filter for grid points within a square, then circular, distance for those closest to the centroid.", echo = FALSE, out.width = "100%", height = 4}

centroids <- centroids %>%
        group_nest(sa2_name_2011) %>%
        mutate(closest = purrr::map(data, closest_focal_point, 
                                    focal_points = capital_cities)) %>%
        unnest(c("data", "closest")) %>%
        arrange(focal_distance) %>% 
        mutate(rownumber = row_number())

f_centroid <- centroids %>%
        filter(rownumber == 1)

# filter_grid_points

distance <- (((f_centroid$latitude - f_centroid$focal_latitude) ^ 2) + 
         ((f_centroid$longitude - f_centroid$focal_longitude) ^ 2)) ^ (1 / 2)

flong <- f_centroid$longitude
flat <- f_centroid$latitude
f_dist <- 10 * hex_size
angle_width <- 35

@@

f1 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) +
  geom_point(aes(x = focal_longitude, y = focal_latitude), data = f_centroid, colour = "red", size = 2) + 
  geom_point(aes(x=longitude, y = latitude), data = f_centroid, colour = "#374926") + 
  theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#9ad5c0", data = grid, size = 0.5) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#64b1be", data = f_grid, size = 2) + ggtitle("a")

f_grid <- f_grid %>%
  group_by(id) %>%
  mutate(
    hex_lat_c = (hex_lat - flat),
    hex_long_c = (hex_long - flong)) %>%
  mutate(hyp = ((hex_lat_c^2) + (hex_long_c^2))^(1 / 2))

f_angle <- f_centroid %>%
      mutate(atan = atan2(latitude - focal_latitude, longitude - focal_longitude),
        angle = (atan * 180 / pi),
        pangle = ifelse(angle < 0, angle + 360, angle)) %>%
      pull(pangle)

f_grid <- f_grid %>%
  # create circle of radius: f_dist
  filter(hyp < f_dist) %>%
  mutate(
    # geosphere takes a long time
    angle = f_angle,
    angle_plus = (angle + angle_width) %% 360,
    angle_minus = (angle - angle_width) %% 360,
    atan = atan2(hex_lat_c, hex_long_c),
    hex_angle = (atan * 180 / pi),
    hex_angle = ifelse(hex_angle < 0, hex_angle + 360, hex_angle))

f2 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) +
  geom_point(aes(x = focal_longitude, y = focal_latitude), data = f_centroid, colour = "red", size = 2) + 
  geom_point(aes(x=longitude, y = latitude), data = f_centroid, colour = "#374926") + 
  theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#9ad5c0", data = grid, size = 0.5) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#64b1be", data = f_grid, size = 2) + ggtitle("b")

grid.arrange(f1,f2, nrow = 1)
```

The `width` parameter is used to take a slice of the remaining points. The width is the amount of degrees used on either side of the angle from the focal point to centroid location. This uses the angle from the closest capital city, to the current centroid as seen in Figure \ref{fig:buffers2} . This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

```{r buffers2, fig.cap = "Filter for grid points within a square, then circular, distance for those closest to the centroid.", echo = FALSE, out.width = "100%", height = 4}

full_grid <- full_grid %>% filter(hex_long > bbox$min[1], hex_long < bbox$max[1],
                                  hex_lat > bbox$min[2], hex_lat < bbox$max[2])

@@

f3 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "lightgrey", colour = "black", size = 0.01) +
  geom_point(aes(x = focal_longitude, y = focal_latitude), data = f_centroid, colour = "red", size = 2) + 
  geom_point(aes(x=longitude, y = latitude), data = f_centroid, colour = "#374926") + 
  theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#9ad5c0", data = grid, size = 0.5) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#64b1be", data = f_grid, size = 2) + ggtitle("a")

>>>>

```{r setup}
library(cartogram)
library(cowplot)
library(gridExtra)
library(ggthemes)
library(knitcitations)
library(kableExtra)
library(sp)
library(sf)
library(sugarbag)
library(tidyverse)

@@

# function to allocate colours to regions
aus_colours <- function(sir_p50){
  value <- case_when(
    sir_p50 <  0.74 ~ "#33809d",
    sir_p50 >= 0.74 & sir_p50 < 0.98 ~ "#aec6c7",
    sir_p50 >= 0.98 & sir_p50 < 1.05 ~ "#fff4bc",
    sir_p50 >= 1.05 & sir_p50 < 1.45 ~ "#ff9a64",
    sir_p50 >= 1.45 ~ "#ff3500",
    TRUE ~ "#FFFFFF")
  return(value)
}
```

@@

```{r thyroiddata}
sa2 <- absmapsdata::sa22011 %>% 
  filter(!st_is_empty(geometry)) %>% 
  filter(!state_name_2011 == "Other Territories") %>% 
  filter(!sa2_name_2011 == "Lord Howe Island")
# sa2 <- sa2 %>% rmapshaper::ms_simplify(keep = 0.05, keep_shapes = TRUE)
SIR <- read_csv("data/SIR Downloadable Data.csv") %>% 
  filter(SA2_name %in% sa2$sa2_name_2011) %>% 
  dplyr::select(Cancer_name, SA2_name, Sex_name, p50) %>% 
  filter(Cancer_name == "Thyroid", Sex_name == "Females")
ERP <- read_csv("data/ERP.csv") %>%
  filter(REGIONTYPE == "SA2", Time == 2011, Region %in% SIR$SA2_name) %>% 
  dplyr::select(Region, Value)
# Alternative maps
# Join with sa2 sf object
sa2thyroid_ERP <- SIR %>% 
  left_join(sa2, ., by = c("sa2_name_2011" = "SA2_name")) %>%
  left_join(., ERP %>% 
              dplyr::select(Region, 
              Population = Value), by = c("sa2_name_2011"= "Region")) %>% 
  filter(!st_is_empty(geometry))
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  #filter(!is.na(Population)) %>% 
  filter(!sa2_name_2011 == "Lord Howe Island") %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf() 
```

@@

```{r fullhexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}
if (!file.exists("data/aus_hexmap.rda")) {

## Create centroids set
centroids <- sa2 %>% 
  create_centroids(., "sa2_name_2011")
## Create hexagon grid
grid <- create_grid(centroids = centroids,
                    hex_size = 0.2,
                    buffer_dist = 5)
## Allocate polygon centroids to hexagon grid points
aus_hexmap <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.2,
  ## same size used in create_grid
  hex_filter = 10,
  focal_points = capital_cities,
  width = 35,
  verbose = FALSE
)
save(aus_hexmap, file = "data/aus_hexmap.rda") }

@@

## Prepare to plot
fort_hex <- fortify_hexagon(data = aus_hexmap,
                            sf_id = "sa2_name_2011",
                            hex_size = 0.2) %>% 
            left_join(sa2thyroid_ERP %>% select(sa2_name_2011, SIR, p50))
fort_aus <- sa2thyroid_ERP %>%
  fortify_sfc()
## Make a plot
aus_hexmap_plot <- ggplot() +
  geom_polygon(aes(x = long,  y = lat,  group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey",  size = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal()
```

@@

# Contiguous Cartograms
if (!file.exists("data/ncont.rda")) {
cont <- sa2thyroid_ERP %>% 
  mutate(Population = Population + 1) %>% 
  cartogram_cont(., weight = "Population", itermax = 20) %>%
  st_as_sf()
save(cont, file = "data/cont.rda") }

@@

tas_ggcont <- ggplot(cont) + 
  geom_sf(fill = "#D3D3D3", colour = "white", data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + guides(fill=FALSE) +
  theme_map() + ggtitle("a")
#tas_ggcont
```

```{r contcart}
# Non - Contiguous Cartograms
if (!file.exists("data/ncont.rda")) {
ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1.5,
                         weight = "Population") %>% 
                          st_as_sf()
save(ncont, file = "data/ncont.rda")
}
```

```{r plotcontcart}
load("data/ncont.rda")
tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "#D3D3D3", colour = "white",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  guides(fill=FALSE) +
  theme_map() + ggtitle("b")
#tas_ggncont
```

```{r dorl}
# Non - Contiguous Dorling Cartograms
dorl <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
tas_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "#D3D3D3", colour = "white",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +
  guides(fill = FALSE) +
  theme_map() + ggtitle("c")
#tas_ggdorl
```

@@

```{r hexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid for males, they range from much lower than average (blue) to much higher than average (red)."}
aus_hexmap_plot
```

@@

```{r tas_centroids, fig.cap = "The Australian Cancer Atlas data has determined the colour of each Statistical Area of Australian at Level 2. A choropleth map (a) of Standardised Incidence Rates (SIRs) is paired with a hexagon tile map (b) to contrast the colours that are made obvious when every SA2 is equally represented."}
centroids <- create_centroids(
  shp_sf = sa2 %>% filter(state_name_2011 == "Tasmania"), 
  sf_id = "sa2_name_2011")
hex_size <- .2
buffer_dist <- 2
bbox <- tibble(min = c(min(centroids$longitude), min(centroids$latitude)),
  max = c(max(centroids$longitude), max(centroids$latitude)))
grid <- tibble::as_tibble(expand.grid(hex_long = seq(bbox$min[1] - buffer_dist,
  bbox$max[1] + buffer_dist,
  hex_size),
  hex_lat = seq(bbox$min[2] - buffer_dist,
    bbox$max[2] + buffer_dist,
    hex_size)))

@@

tas_centroids <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "#D3D3D3", colour = "white", size = 0.01) + 
  geom_point(aes(x=longitude, y = latitude), size = 2, shape = 19, colour = "white",
             data = centroids) +
  scale_colour_identity() +
  theme_void() + 
  coord_equal()

tas_hexmap <- ggplot() +
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "#D3D3D3", colour = "white", size = 0.01) + 
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR),
               data = fort_hex %>% filter(sa2_name_2011 %in% fort_tas$sa2_name_2011)) +
  scale_fill_identity() +
  theme_void() + 
  coord_equal()
```

@@

@@

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is the size specified by `hex_size`. Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. Similarly, the rows are created from the latitude values and the buffer distance. A unique hexagon location is created from all intersections of the longitude columns and latitude rows. Figure \ref{fig:filterprocess} shows the original grid on the left, to allow for tessellating hexagons, every second latitude row on the grid is shifted right, by half of the hexagon size. The grid for tessellation is shown on the right.

```{r grid, fig.cap = "Complete set of grid points to create a tile map.", height = 4}
# Find every second latitude
shift_lat <- grid %>% dplyr::select(hex_lat) %>%
  dplyr::distinct() %>%
  dplyr::filter(dplyr::row_number() %% 2 == 1) %>% unlist()
# Shift the longitude of every second latitude to the right to make hex structure
grid <- grid %>%
  dplyr::mutate(hex_long = ifelse(hex_lat %in% shift_lat, hex_long,
    hex_long + (hex_size / 2))) %>%
  dplyr::mutate(id=1:NROW(.))  %>%
  dplyr::mutate(assigned=FALSE)
```

@@

The closest latitude row and longitude column are found for each centroid location. Then rows and columns of centroids are divided into 20 groups. The amount of rows in each latitude group and the amount of columns in each longitude group are used as the width of rolling windows. The rolling windows can be seen on the bottom and right of the grid shown in Figure \ref{fig:filterprocess}. This will tailor the available grid points to those most likely to be used. It also helps reduce the amount of time taken, as it decreases the amount of points considered for each centroid allocation.

@@

```{r filtergrid, fig.cap = "The remaining hexagon locations from all the possible points in the initial grid after the buffer is applied. The blue dots show the grid points left to choose from after the buffer step."}
g1 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "#D3D3D3", colour = "white", size = 0.01) + 
   theme_void() + 
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", data = grid, size = 1) +
  coord_equal() + ggtitle("a")

full_grid <- grid <- grid %>%
  mutate(hex_long_int = dense_rank(hex_long)-1,
    hex_lat_int = dense_rank(hex_lat)-1)
nlong <- length(unique(grid$hex_long))
nlat <- length(unique(grid$hex_lat))

# Add grid row and column values to centroids data set 
centroids <- centroids %>%
  mutate(
    # long int for integer value of longitude column
    long_int = round((longitude - min(grid$hex_long)) / 
                       (max(grid$hex_long) - min(grid$hex_long)) * nlong, 0),
    # lat int for integer value of latitude column
    lat_int = round((latitude - min(grid$hex_lat)) / 
                      (max(grid$hex_lat) - min(grid$hex_lat)) * nlat, 0)
  )
# Amount of lats and longs in each group of rows and columns
lat_size <- round(nlat / 20, 0)
long_size <- round(nlong / 20, 0)
# Make a list of the min and max of the groups
# Effectively creates manual sliding windows
nlat_list <- purrr::map2(seq(1:nlat), lat_size + seq(1:nlat), c)
nlong_list <- purrr::map2(seq(1:nlong), long_size + seq(1:nlong), c)
# LATITUDE ROWS FILTER

@@

  max_int <- min(bounds[2], maximum)
  cents_in <- filter(cents, between(lat_int, bounds[1], max_int))
  return(cents_in)
}

@@

  while (max_int > maximum) {
    max_int <- max_int - 1
  }
  cents_in <- filter(cents, between(long_int, bounds[1], max_int))
  return(cents_in)
}

@@

# DEFINE FUNCTION
# find the min and max longitude for each latitude
range_rows <- purrr::map_dfr(
  .x = lat_windows,
  .f = function(data) {
    data %>%
      dplyr::summarise(
        long_min = ifelse(purrr::is_empty(long_int), NA, min(data$long_int)),
        long_max = ifelse(purrr::is_empty(long_int), NA, max(data$long_int))
      )
  }
)
# find the min and max longitude for each latitude
range_cols <- purrr::map_dfr(.x = long_windows, .f = function(data) {
  data %>%
    dplyr::summarise(
      lat_min = ifelse(purrr::is_empty(lat_int), NA, min(data$lat_int)),
      lat_max = ifelse(purrr::is_empty(lat_int), NA, max(data$lat_int))
    )
})

@@

# smooth the minimums
av_range_cols <- purrr::map_dfr(.x = nlong_list, mean_range, range_cols) %>%
  bind_cols(long_id = c(seq(1:nlong) + round(long_size / 2)), .)
# APPLY A BUFFER
# change buffer to amount of hexagons (ints) either side
hex_buffer <- floor(buffer_dist / hex_size)

@@

g2 <- ggplot() + 
   geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "#D3D3D3", colour = "white", size = 0.01) +  
  theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#55b56c", data = full_grid, size = 0.5) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", data = grid, size = 1.5) + ggtitle("b")

@@

```{r buffers, fig.cap = "Filter for grid points within a square, then circular, distance for those closest to the centroid.", echo = FALSE, out.width = "100%", height = 4}
centroids <- centroids %>%
        group_nest(sa2_name_2011) %>%
        mutate(closest = purrr::map(data, closest_focal_point, 
                                    focal_points = capital_cities)) %>%
        unnest(c("data", "closest")) %>%
        arrange(focal_distance) %>% 
        mutate(rownumber = row_number())

# Choose a centroid for allocation:
f_centroid <- centroids %>%
        filter(rownumber == 7)

# filter_grid_points
distance <- (((f_centroid$latitude - f_centroid$focal_latitude) ^ 2) + 
         ((f_centroid$longitude - f_centroid$focal_longitude) ^ 2)) ^ (1 / 2)
flong <- f_centroid$longitude
flat <- f_centroid$latitude
f_dist <- 10 * hex_size
angle_width <- 35

@@

f_grid <- f_grid %>%
  group_by(id) %>%
  mutate(
    hex_lat_c = (hex_lat - flat),
    hex_long_c = (hex_long - flong)) %>%
  mutate(hyp = ((hex_lat_c^2) + (hex_long_c^2))^(1 / 2))
f_angle <- f_centroid %>%
      mutate(atan = atan2(latitude - focal_latitude, longitude - focal_longitude),
        angle = (atan * 180 / pi),
        pangle = ifelse(angle < 0, angle + 360, angle)) %>%
      pull(pangle)
f_grid <- f_grid %>%
  # create circle of radius: f_dist
  filter(hyp < f_dist) %>%
  mutate(
    # geosphere takes a long time
    angle = f_angle,
    angle_plus = (angle + angle_width) %% 360,
    angle_minus = (angle - angle_width) %% 360,
    atan = atan2(hex_lat_c, hex_long_c),
    hex_angle = (atan * 180 / pi),
    hex_angle = ifelse(hex_angle < 0, hex_angle + 360, hex_angle))

f1 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "#D3D3D3", colour = "white", size = 0.01) +
  theme_void() + coord_equal() +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#55b56c", data = grid, size = 0.5) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", data = f_grid, size = 1.5) + 
  geom_point(aes(x=longitude, y = latitude), data = f_centroid, fill = "#ede187", shape = 24) + 
#Hobart
  geom_point(aes(x = focal_longitude, y = focal_latitude), data = f_centroid, colour = "#e9f2ed", size = 2, shape = 3) + ggtitle("c")

The `width` parameter is used to take a slice of the remaining points. The width is the amount of degrees used on either side of the angle from the focal point to centroid location. This uses the angle from the closest capital city, to the current centroid as seen in Figure \ref{fig:filterprocess} . This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

```{r filterprocess, fig.cap = "Filter for grid points within a square, then circular, distance for those closest to the centroid.", echo = FALSE, out.width = "120%", height = 8}
full_grid <- full_grid %>% filter(hex_long > bbox$min[1], hex_long < bbox$max[1],
                                  hex_lat > bbox$min[2], hex_lat < bbox$max[2])

allocated_list <- aus_hexmap %>% 
                  filter(points == "Hobart") %>% 
                  slice_min(rownumber, n = 6)

fort_allocated <- allocated_list %>% 
  fortify_hexagon(., sf_id = "sa2_name_2011", hex_size = 0.2)

@@
